---
title: "Simulating SMART designs with sim_treatment"
subtitle: "Building two-stage SMARTs step by step"
output: 
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulating SMART designs with sim_treatment}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4
)
```

## Introduction

The `sim_treatment()` function provides a flexible way to simulate treatment
assignments one stage at a time.
It can be used to simulate various common SMART designs, including those with response indicators or feasible sets. 
An additional function `block_rand()` can implement permuted block randomization to ensure greater balance than binomial randomization in later stages.

This vignette walks through three two-stage designs:

1. **No response status** — all participants are re-randomized at stage 2.
2. **Response, non-responders only** — a response indicator is observed
   between stages and only non-responders are re-randomized.
3. **Response, everyone re-randomized** — a response indicator is observed
   but both responders and non-responders receive a new randomization at
   stage 2.

The first stage helps to build our intuition in the simplest case. The second and third showcase additional features of the function. 
Each section includes a diagram of the design, sample simulation code, and
the resulting data.

```{r load}
library(rsmart)
```

## Design 1: No response status

In the simplest two-stage SMART, every participant is randomized at stage 1
and again at stage 2 regardless of any intermediate outcome.
This design has four embedded treatment regimes.

```{r design1-diagram, echo=FALSE, fig.cap="Design 1: Two-stage SMART without a response. All participants are re-randomized at stage 2.", fig.height=4, fig.width=8}
par(mar = c(1.5, 0.2, 0.5, 0.2), family = "sans")
plot.new()
plot.window(xlim = c(0, 11), ylim = c(-0.8, 6.5))

# Colors (matching reference image)
col_rand <- "#5B7FBA"   # blue circles for randomization
col_trt  <- "#1A8A7A"   # teal boxes for treatments
col_line <- "grey30"    # connecting lines
col_phase <- "#6EC6B8"  # light teal for phase boxes

# Helper: draw a filled rounded-rect label (simulated with rect + text)
draw_box <- function(x, y, label, w = 0.9, h = 0.4, bg = col_trt, tcol = "white", cex = 0.72) {
  rect(x - w, y - h, x + w, y + h, col = bg, border = NA)
  text(x, y, label, cex = cex, col = tcol, font = 2)
}

# --- Stage 1 randomization node (circle) ---
symbols(1, 3, circles = 0.35, inches = FALSE, add = TRUE,
        bg = col_rand, fg = col_rand)
text(1, 3, "R", cex = 0.95, font = 2, col = "white")

# --- Stage 1 branches ---
segments(1.35, 3, 2.8, 5, lwd = 2, col = col_line)
segments(1.35, 3, 2.8, 1, lwd = 2, col = col_line)

# --- Stage 1 treatment boxes ---
draw_box(3.7, 5, "Treatment A")
draw_box(3.7, 1, "Treatment B")
segments(2.8, 5, 2.8, 5, lwd = 2, col = col_line)  # connector
segments(2.8, 1, 2.8, 1, lwd = 2, col = col_line)

# Arrows from branch end to box
arrows(2.8, 5, 2.8, 5, length = 0, lwd = 2, col = col_line)
segments(2.8, 5, 3.7 - 0.9, 5, lwd = 2, col = col_line)
segments(2.8, 1, 3.7 - 0.9, 1, lwd = 2, col = col_line)

# --- Connectors from treatment boxes to stage 2 R nodes ---
segments(3.7 + 0.9, 5, 5.8 - 0.35, 5, lwd = 2, col = col_line)
segments(3.7 + 0.9, 1, 5.8 - 0.35, 1, lwd = 2, col = col_line)

# --- Stage 2 randomization nodes (circles) ---
symbols(5.8, 5, circles = 0.35, inches = FALSE, add = TRUE,
        bg = col_rand, fg = col_rand)
text(5.8, 5, "R", cex = 0.85, font = 2, col = "white")

symbols(5.8, 1, circles = 0.35, inches = FALSE, add = TRUE,
        bg = col_rand, fg = col_rand)
text(5.8, 1, "R", cex = 0.85, font = 2, col = "white")

# --- Stage 2 branches (upper: a1=0) ---
segments(6.15, 5, 7.5, 5.9, lwd = 1.8, col = col_line)
segments(6.15, 5, 7.5, 4.1, lwd = 1.8, col = col_line)

# --- Stage 2 branches (lower: a1=1) ---
segments(6.15, 1, 7.5, 1.9, lwd = 1.8, col = col_line)
segments(6.15, 1, 7.5, 0.1, lwd = 1.8, col = col_line)

# Connectors to outcome boxes
segments(7.5, 5.9, 9.5 - 0.9, 5.9, lwd = 1.8, col = col_line)
segments(7.5, 4.1, 9.5 - 0.9, 4.1, lwd = 1.8, col = col_line)
segments(7.5, 1.9, 9.5 - 0.9, 1.9, lwd = 1.8, col = col_line)
segments(7.5, 0.1, 9.5 - 0.9, 0.1, lwd = 1.8, col = col_line)

# --- Outcome treatment boxes ---
draw_box(9.5, 5.9, "Treatment C", w = 0.95)
draw_box(9.5, 4.1, "Treatment D", w = 0.95)
draw_box(9.5, 1.9, "Treatment E", w = 0.95)
draw_box(9.5, 0.1, "Treatment F", w = 0.95)

# --- Stage legend at the bottom ---
draw_box(x = 1, y = -0.6, label = "Phase:", 
        w = 0.5, h = 0.18, bg = col_phase, tcol = "white", cex = 0.7)

draw_box(x = 3.7, y = -0.6, label = "Stage 1", 
         h = 0.18, bg = col_phase, tcol = "white", cex = 0.7)

draw_box(x = 9.5, y = -0.6, label = "Stage 2", 
         h = 0.18, bg = col_phase, tcol = "white", cex = 0.7)
```

This design has four embedded regimes: 
(A, C), (A, D), (B, E), and (B, F), which will be encoded as
`(a1=0, a2=0)`, `(a1=0, a2=1)`, `(a1=1, a2=0)`,
`(a1=1, a2=1)`, respectively, using `sim_treatment()`. 

### Simulation

Consider a possible trial with `500` participants. 
The age of participants is recorded at entry and their compliance with treatment at stage 2. 
We assign treatments with equal randomization using a permuted block design. 

We provide `sim_treatment()` with:

- `n_treatments`: the number of treatments available at the stage for randomized participants
- `dat`: a dataset with the initial covariates and IDs at the first stage
- `stage`: the stage of the trial that treatments should be assigned for
- `rand_prob_fn`: the `block_rand` function to generate permuted block randomization, with each treatment repeated twice in a block

The `block_rand` function here will use a block size of 4 so a sample block could be `2 1 1 2`. 
This will ensure that treatment imbalance between groups is minimized.

Because there is no response status, the input `randomize_response` is implicitly left at its
default (`NULL`).

The function will return the data frame with the treatment assignment for each participant.

```{r design1-sim}
set.seed(42)
n <- 500

# Start with baseline data
dat1 <- data.frame(
  id  = 1:n,
  x1  = round(runif(n, 25, 75)),        # age
  x2  = rbinom(n, 1, 0.9)               # treatment compliance
)

# Stage 1: block randomization
dat1 <- sim_treatment(n_treatments = 2, dat = dat1, stage = 1,
                      rand_prob_fn = block_rand(block_rep = 2))

# Stage 2: everyone is re-randomized within a1 groups (no response)
dat1 <- sim_treatment(n_treatments = 2, dat = dat1, stage = 2,
                      rand_prob_fn = block_rand(block_rep = 2))

table(dat1$a1, dat1$a2)
head(dat1)
```

We see that the resulting data set has the columns `r colnames(dat1)`. 
The treatments are appended as `a<stage>`.  

## Design 2: Only non-responders are re-randomized

Commonly, tailoring variables are used in SMARTs to determine sequential treatment options. 
A response indicator is observed between stages and then governs available treatements.
Often, it may be unethical to re-randomized away from a treatment that is working well and 
responders are given a second stage treatment deterministically. 
Alternatively, non-responders may be discontinued or given a single salvage therapy. 
Here we encode this status using `r2` and will assign responders `r2 = 1` a single treatment 
(coded `a2 = 0`).
Non-responders will be re-randomized. 

In the case where responders are randomized and non-responders are not, then the encoding of `r2 = 1` represents 
non-responders and the function `sim_treatment` will assign responders, now encoded as `r2 = 0`, treatments. 

```{r design2-diagram, echo=FALSE, fig.cap="Design 2: Two-stage SMART where only non-responders are re-randomized at stage 2.", fig.height=5.5, fig.width=9}
par(mar = c(1.5, 0.2, 0.5, 0.2), family = "sans")
plot.new()
plot.window(xlim = c(0, 13), ylim = c(-1.5, 9))

# Colors
col_rand  <- "#5B7FBA"
col_trt   <- "#1A8A7A"
col_line  <- "grey30"
col_resp  <- "grey50"
col_phase <- "#6EC6B8"

draw_box <- function(x, y, label, w = 1.0, h = 0.4, bg = col_trt, tcol = "white", cex = 0.7) {
  rect(x - w, y - h, x + w, y + h, col = bg, border = NA)
  text(x, y, label, cex = cex, col = tcol, font = 2)
}

# --- Stage 1 randomization node ---
symbols(1, 4, circles = 0.38, inches = FALSE, add = TRUE,
        bg = col_rand, fg = col_rand)
text(1, 4, "R", cex = 0.95, font = 2, col = "white")

# --- Stage 1 branches ---
segments(1.38, 4, 2.5, 7, lwd = 2, col = col_line)
segments(1.38, 4, 2.5, 1, lwd = 2, col = col_line)

# --- Stage 1 treatment boxes ---
draw_box(3.5, 7, "Treatment A")
draw_box(3.5, 1, "Treatment B")
segments(2.5, 7, 3.5 - 1.0, 7, lwd = 2, col = col_line)
segments(2.5, 1, 3.5 - 1.0, 1, lwd = 2, col = col_line)

# --- Response decision boxes ---
draw_box(5.7, 7, "Response?", w = 0.9, bg = col_phase)
draw_box(5.7, 1, "Response?", w = 0.9, bg = col_phase)
segments(3.5 + 1.0, 7, 5.7 - 0.9, 7, lwd = 2, col = col_line)
segments(3.5 + 1.0, 1, 5.7 - 0.9, 1, lwd = 2, col = col_line)

# --- Upper arm (a1 = 0) response branches ---
# Yes (responder) - dashed
segments(6.6, 7, 7.5, 8.2, lwd = 1.5, col = col_resp, lty = 2)
text(7.2, 8.5, "Yes", cex = 0.7, col = col_resp, font = 3)
# No (non-responder) - solid
segments(6.6, 7, 7.5, 5.8, lwd = 1.8, col = col_line)
text(7.2, 5.5, "No", cex = 0.7, col = col_line, font = 3)

# Responder deterministic treatment
segments(7.5, 8.2, 10.8 - 1.0, 8.2, lwd = 1.5, col = col_resp, lty = 2)
draw_box(10.8, 8.2, "Treatment C", w = 1.0)

# Non-responder randomization
symbols(8.2, 5.8, circles = 0.32, inches = FALSE, add = TRUE,
        bg = col_rand, fg = col_rand)
text(8.2, 5.8, "R", cex = 0.75, font = 2, col = "white")
segments(7.5, 5.8, 8.2 - 0.32, 5.8, lwd = 1.8, col = col_line)

# Non-responder branches (upper)
segments(8.52, 5.8, 9.3, 6.6, lwd = 1.5, col = col_line)
segments(8.52, 5.8, 9.3, 5.0, lwd = 1.5, col = col_line)
segments(9.3, 6.6, 10.8 - 1.0, 6.6, lwd = 1.5, col = col_line)
segments(9.3, 5.0, 10.8 - 1.0, 5.0, lwd = 1.5, col = col_line)
draw_box(10.8, 6.6, "Treatment D", w = 1.0)
draw_box(10.8, 5.0, "Treatment E", w = 1.0)

# --- Lower arm (a1 = 1) response branches ---
# Yes (responder) - dashed
segments(6.6, 1, 7.5, 2.2, lwd = 1.5, col = col_resp, lty = 2)
text(7.2, 2.5, "Yes", cex = 0.7, col = col_resp, font = 3)
# No (non-responder) - solid
segments(6.6, 1, 7.5, -0.2, lwd = 1.8, col = col_line)
text(7.2, -0.5, "No", cex = 0.7, col = col_line, font = 3)

# Responder deterministic treatment
segments(7.5, 2.2, 10.8 - 1.0, 2.2, lwd = 1.5, col = col_resp, lty = 2)
draw_box(10.8, 2.2, "Treatment F", w = 1.0)

# Non-responder randomization
symbols(8.2, -0.2, circles = 0.32, inches = FALSE, add = TRUE,
        bg = col_rand, fg = col_rand)
text(8.2, -0.2, "R", cex = 0.75, font = 2, col = "white")
segments(7.5, -0.2, 8.2 - 0.32, -0.2, lwd = 1.8, col = col_line)

# Non-responder branches (lower)
segments(8.52, -0.2, 9.3, 0.6, lwd = 1.5, col = col_line)
segments(8.52, -0.2, 9.3, -1.0, lwd = 1.5, col = col_line)
segments(9.3, 0.6, 10.8 - 1.0, 0.6, lwd = 1.5, col = col_line)
segments(9.3, -1.0, 10.8 - 1.0, -1.0, lwd = 1.5, col = col_line)
draw_box(10.8, 0.6, "Treatment G", w = 1.0)
draw_box(10.8, -1.0, "Treatment H", w = 1.0)

# --- Phase legend ---
draw_box(x = 1, y = -1.7, label = "Phase:",
         w = 0.5, h = 0.18, bg = col_phase, tcol = "white", cex = 0.65)
draw_box(x = 3.5, y = -1.7, label = "Stage 1",
         w = 0.9, h = 0.18, bg = col_phase, tcol = "white", cex = 0.65)
draw_box(x = 5.7, y = -1.7, label = "Response",
         w = 0.9, h = 0.18, bg = col_phase, tcol = "white", cex = 0.65)
draw_box(x = 10.8, y = -1.7, label = "Stage 2",
         w = 0.9, h = 0.18, bg = col_phase, tcol = "white", cex = 0.65)
```

Responders follow a deterministic path (dashed lines). 
There are still four embedded regimes, but now follow rules 
"give treamtent 1 at stage 1, if response give treatment 2 at stage 2, 
otherwise give treatment 3 at stage 2. (1, 2, 3)"
The embedded regimes are now encoded as:
`(a1=0, a2=0, a2=0)`, 
`(a1=0, a2=0, a2=1)`, 
`(a1=1, a2=0, a2=0)`, 
`(a1=1, a2=0, a2=1)`.

### Simulation

We use similar inputs as before, but with the additional input 
`randomize_response = "N"` so that we assign `a2 = 0` to all responders and
randomize only the non-responders within each `a1` group.

```{r design2-sim}
set.seed(42)
n <- 500

dat2 <- data.frame(
  id  = 1:n,
  t1  = sort(round(runif(n, 0, 365 * 2))),  # enrollment day
  x1  = round(runif(n, 25, 75))             # age
)

# Stage 1: block randomization (data sorted by t1)
dat2 <- sim_treatment(n_treatments = 2, dat = dat2, stage = 1,
                      rand_prob_fn = block_rand(block_rep = 2))

# Generate response and stage 2 timing
dat2$r2 <- rbinom(n, 1, 0.5)
dat2$t2 <- dat2$t1 + 100 + round(runif(n, -10, 10))

# Sort by t2 before stage 2 randomization (see note below)
dat2 <- dat2[order(dat2$t2), ]

# Stage 2: only non-responders re-randomized
dat2 <- sim_treatment(n_treatments = 2, dat = dat2, stage = 2,
                      rand_prob_fn = block_rand(block_rep = 2),
                      randomize_response = "N")

table(dat2$a1, dat2$a2, dat2$r2,
      dnn = c("a1", "a2", "r2"))
```

Because of the use of the permuted block randomization, we see that among responders and conditioned on the first treatment, 
there is relative balance between treatment assignments.
All responders `r2 = 1` received `a2 = 0`. 


## Design 3: Both responders and non-responders re-randomized

In some SMARTs, all participants are re-randomized at stage 2, with treatments allowed to vary
by response status. 
This means responders may or may not receive different second-stage treatments than
non-responders, and the randomization probabilities or allocation schedules
can differ between the two groups.

```{r design3-diagram, echo=FALSE, fig.cap="Design 3: Two-stage SMART where both responders and non-responders are re-randomized at stage 2.", fig.height=6.5, fig.width=9}
par(mar = c(1.5, 0.2, 0.5, 0.2), family = "sans")
plot.new()
plot.window(xlim = c(0, 13), ylim = c(-2.5, 10.5))

# Colors
col_rand  <- "#5B7FBA"
col_trt   <- "#1A8A7A"
col_line  <- "grey30"
col_resp  <- "grey50"
col_phase <- "#6EC6B8"

draw_box <- function(x, y, label, w = 1.0, h = 0.4, bg = col_trt, tcol = "white", cex = 0.7) {
  rect(x - w, y - h, x + w, y + h, col = bg, border = NA)
  text(x, y, label, cex = cex, col = tcol, font = 2)
}

# --- Stage 1 randomization node ---
symbols(1, 4, circles = 0.38, inches = FALSE, add = TRUE,
        bg = col_rand, fg = col_rand)
text(1, 4, "R", cex = 0.95, font = 2, col = "white")

# --- Stage 1 branches ---
segments(1.38, 4, 2.5, 7.5, lwd = 2, col = col_line)
segments(1.38, 4, 2.5, 0.5, lwd = 2, col = col_line)

# --- Stage 1 treatment boxes ---
draw_box(3.5, 7.5, "Treatment A")
draw_box(3.5, 0.5, "Treatment B")
segments(2.5, 7.5, 3.5 - 1.0, 7.5, lwd = 2, col = col_line)
segments(2.5, 0.5, 3.5 - 1.0, 0.5, lwd = 2, col = col_line)

# --- Response decision boxes ---
draw_box(5.7, 7.5, "Response?", w = 0.9, bg = col_phase)
draw_box(5.7, 0.5, "Response?", w = 0.9, bg = col_phase)
segments(3.5 + 1.0, 7.5, 5.7 - 0.9, 7.5, lwd = 2, col = col_line)
segments(3.5 + 1.0, 0.5, 5.7 - 0.9, 0.5, lwd = 2, col = col_line)

# ===== Upper arm (a1 = 0) =====
# Yes (responder)
segments(6.6, 7.5, 7.3, 9.2, lwd = 1.8, col = col_line)
text(7.0, 9.5, "Yes", cex = 0.7, col = col_resp, font = 3)
# No (non-responder)
segments(6.6, 7.5, 7.3, 5.8, lwd = 1.8, col = col_line)
text(7.0, 5.5, "No", cex = 0.7, col = col_line, font = 3)

# Responder randomization (a1=0, resp)
symbols(7.9, 9.2, circles = 0.30, inches = FALSE, add = TRUE,
        bg = col_rand, fg = col_rand)
text(7.9, 9.2, "R", cex = 0.65, font = 2, col = "white")
segments(7.3, 9.2, 7.9 - 0.30, 9.2, lwd = 1.5, col = col_line)

segments(8.2, 9.2, 9.0, 9.9, lwd = 1.3, col = col_line)
segments(8.2, 9.2, 9.0, 8.5, lwd = 1.3, col = col_line)
segments(9.0, 9.9, 10.8 - 1.0, 9.9, lwd = 1.3, col = col_line)
segments(9.0, 8.5, 10.8 - 1.0, 8.5, lwd = 1.3, col = col_line)
draw_box(10.8, 9.9, "Treatment C", w = 1.0)
draw_box(10.8, 8.5, "Treatment D", w = 1.0)

# Non-responder randomization (a1=0, non-resp)
symbols(7.9, 5.8, circles = 0.30, inches = FALSE, add = TRUE,
        bg = col_rand, fg = col_rand)
text(7.9, 5.8, "R", cex = 0.65, font = 2, col = "white")
segments(7.3, 5.8, 7.9 - 0.30, 5.8, lwd = 1.5, col = col_line)

segments(8.2, 5.8, 9.0, 6.5, lwd = 1.3, col = col_line)
segments(8.2, 5.8, 9.0, 5.1, lwd = 1.3, col = col_line)
segments(9.0, 6.5, 10.8 - 1.0, 6.5, lwd = 1.3, col = col_line)
segments(9.0, 5.1, 10.8 - 1.0, 5.1, lwd = 1.3, col = col_line)
draw_box(10.8, 6.5, "Treatment E", w = 1.0)
draw_box(10.8, 5.1, "Treatment F", w = 1.0)

# ===== Lower arm (a1 = 1) =====
# Yes (responder)
segments(6.6, 0.5, 7.3, 2.2, lwd = 1.8, col = col_line)
text(7.0, 2.5, "Yes", cex = 0.7, col = col_resp, font = 3)
# No (non-responder)
segments(6.6, 0.5, 7.3, -1.2, lwd = 1.8, col = col_line)
text(7.0, -1.5, "No", cex = 0.7, col = col_line, font = 3)

# Responder randomization (a1=1, resp)
symbols(7.9, 2.2, circles = 0.30, inches = FALSE, add = TRUE,
        bg = col_rand, fg = col_rand)
text(7.9, 2.2, "R", cex = 0.65, font = 2, col = "white")
segments(7.3, 2.2, 7.9 - 0.30, 2.2, lwd = 1.5, col = col_line)

segments(8.2, 2.2, 9.0, 2.9, lwd = 1.3, col = col_line)
segments(8.2, 2.2, 9.0, 1.5, lwd = 1.3, col = col_line)
segments(9.0, 2.9, 10.8 - 1.0, 2.9, lwd = 1.3, col = col_line)
segments(9.0, 1.5, 10.8 - 1.0, 1.5, lwd = 1.3, col = col_line)
draw_box(10.8, 2.9, "Treatment G", w = 1.0)
draw_box(10.8, 1.5, "Treatment H", w = 1.0)

# Non-responder randomization (a1=1, non-resp)
symbols(7.9, -1.2, circles = 0.30, inches = FALSE, add = TRUE,
        bg = col_rand, fg = col_rand)
text(7.9, -1.2, "R", cex = 0.65, font = 2, col = "white")
segments(7.3, -1.2, 7.9 - 0.30, -1.2, lwd = 1.5, col = col_line)

segments(8.2, -1.2, 9.0, -0.5, lwd = 1.3, col = col_line)
segments(8.2, -1.2, 9.0, -1.9, lwd = 1.3, col = col_line)
segments(9.0, -0.5, 10.8 - 1.0, -0.5, lwd = 1.3, col = col_line)
segments(9.0, -1.9, 10.8 - 1.0, -1.9, lwd = 1.3, col = col_line)
draw_box(10.8, -0.5, "Treatment I", w = 1.0)
draw_box(10.8, -1.9, "Treatment J", w = 1.0)

# --- Phase legend ---
draw_box(x = 1, y = -2.6, label = "Phase:",
         w = 0.5, h = 0.18, bg = col_phase, tcol = "white", cex = 0.65)
draw_box(x = 3.5, y = -2.6, label = "Stage 1",
         w = 0.9, h = 0.18, bg = col_phase, tcol = "white", cex = 0.65)
draw_box(x = 5.7, y = -2.6, label = "Response",
         w = 0.9, h = 0.18, bg = col_phase, tcol = "white", cex = 0.65)
draw_box(x = 10.8, y = -2.6, label = "Stage 2",
         w = 0.9, h = 0.18, bg = col_phase, tcol = "white", cex = 0.65)
```

Because both responders and non-responders are randomized at stage 2,
the number of embedded regimes grows.
With the canonical encoding, each regime specifies a stage-1 treatment, a
responder stage-2 treatment, and a non-responder stage-2 treatment, yielding
$2 \times 2 \times 2 = 8$ embedded regimes.
For completeness, we can enumerate these regimes as follows:

1. (0, 0, 0)
2. (0, 0, 1)
3. (0, 1, 0)
4. (0, 1, 1)
5. (1, 0, 0)
6. (1, 0, 1)
7. (1, 1, 0)
8. (1, 1, 1)

Again, using the same format where these are the treatments given at stage 1, 
at stage 2 for responders, and at stage 2 for non-responders. 


### Simulation

Setting `randomize_response = "Y"` includes the response column `r2` in the
grouping variables so that randomization is performed independently within
each combination of `a1` and `r2`.

```{r design3-sim}
set.seed(42)
n <- 500

dat3 <- data.frame(
  id  = 1:n,
  t1  = sort(round(runif(n, 0, 365 * 2))),
  x1  = round(runif(n, 25, 75))
)

# Stage 1
dat3 <- sim_treatment(n_treatments = 2, dat = dat3, stage = 1,
                      rand_prob_fn = block_rand(block_rep = 2))

# Generate response and timing
dat3$r2 <- rbinom(n, 1, 0.5)
dat3$t2 <- dat3$t1 + 100 + round(runif(n, -10, 10))

# Sort by t2 before stage 2
dat3 <- dat3[order(dat3$t2), ]

# Stage 2: responders AND non-responders randomized (stratified by response)
dat3 <- sim_treatment(n_treatments = 2, dat = dat3, stage = 2,
                      rand_prob_fn = block_rand(block_rep = 2),
                      randomize_response = "Y")

table(dat3$a1, dat3$a2, dat3$r2,
      dnn = c("a1", "a2", "r2"))
```

Unlike Design 2, responders now receive both `a2 = 0` and `a2 = 1`.
We again see that conditioned on response status and first treatment, the a2 assignments are relatively well-balanced. 

## Note on ordering and block randomization

When using `block_rand()` (or any allocation-schedule–based randomization),
the order in which participants appear in the data frame determines their
position in the allocation schedule.
If enrollment timing matters — for example, in a group sequential design — the
data should be **sorted by the relevant time variable before calling
`sim_treatment()`** at each stage.

For stage 1, sort by the enrollment time (e.g., `t1`):

```r
dat <- dat[order(dat$t1), ]
dat <- sim_treatment(n_treatments = 2, dat = dat, stage = 1,
                     rand_prob_fn = block_rand(block_rep = 2))
```

For stage 2 and beyond, sort by the time of re-randomization (e.g., `t2`):

```r
dat <- dat[order(dat$t2), ]
dat <- sim_treatment(n_treatments = 2, dat = dat, stage = 2,
                     rand_prob_fn = block_rand(block_rep = 2),
                     randomize_response = "N")
```

This ensures that the permuted block allocation schedule aligns with the
temporal order of participant arrivals at each decision point.
If simple randomization (the default `rand_prob_fn`) is used instead of
block randomization, the row order does not affect the allocation and
sorting is not required.
